### 2022-01-27 _posts/welcome/welcome.Rmd
Add Distill skeleton

---
title: "Welcome to Steve Linberg"
description: |
  Welcome to our new blog, Steve Linberg. We hope you enjoy 
  reading what we have to say!
author:
  - name: Nora Jones 
    url: https://example.com/norajones
    affiliation: Spacely Sprockets
    affiliation_url: https://example.com/spacelysprokets
date: 01-27-2022
output:
  distill::distill_article:
    self_contained: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)

# Learn more about creating blogs with Distill at:
# https://rstudio.github.io/distill/blog.html

```