### 2022-01-28 _posts/welcome/welcome.Rmd
Edit default welcome post

---
title: "New blog under construction"
description: |
  We're re-launching slinberg.net using Distill. Stand by!
author:
  - name: Steve Linberg
    url: https://slinberg.net
    # affiliation: Spacely Sprockets
    # affiliation_url: https://example.com/spacelysprokets
date: 01-27-2022
output:
  distill::distill_article:
    self_contained: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)

# Learn more about creating blogs with Distill at:
# https://rstudio.github.io/distill/blog.html

```

This is a re-launch of slinberg.net, using the [Distill](https://rstudio.github.io/distill) platform to align with the UMASS DACSS program I am presently enrolled in.