### 2022-01-29 _meta/distill-and-github-pages-setup/distill-and-github-pages-setup.Rmd
build and dirs

---
title: "Distill and Github Pages Setup"
description: |
  A short description of the post.
author:
  - name: Steve Linberg
date: 2022-01-29
output:
  distill::distill_article:
    self_contained: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```