### 2022-01-29 steve.Rmd
Add postcard page

---
title: "Steve Linberg"
image: "2021-sfl-photo-fadedbg-256.png"
links:
  - label: LinkedIn
    url: "https://www.linkedin.com/in/slinberg"
  - label: Twitter
    url: "https://twitter.com/stevelinberg"
  - label: GitHub
    url: "https://github.com/stevelinberg"
  - label: SO
    url: "https://stackoverflow.com/users/13603796/stevelinberg"
  - label: Email
    url: "mailto:steve@slinberg.net"
output:
  postcards::trestles
---

## Bio

I am a Masters student in the [DACSS program](https://www.umass.edu/sbs/data-analytics-and-computational-social-science-program) at the University of Massachusetts, Amherst; anticipated graduation Summer 2022. The day job is Senior Systems and Analytics Engineer at the [Gandara Center](https://www.gandaracenter.org/) in West Springfield, MA.

This site contains my academic [blog](index.html) with ruminations on computational social science, my own learning processes and discoveries, and related technical topics and snippets.

## Education

**Hampshire College** | Amherst, MA

BA, Social Science, 1990-3

**University of Massachusetts, Amherst** | Amherst, MA

M.S, Data Analytics and Computational Social Science | Anticipated graduation summer 2022

## Experience

**Gandara Center** | Senior Systems and Analytics Engineer / CIO / IT Director | 2005 - Present

**Silicon Goblin Technologies** | Owner / Programmer | 1999 - 2015

**National Center on Adult Literacy**, University of Pennsylvania Graduate School of Education | Systems Programmer | 1997-1999

**Holyoke Public Schools** | Literacy and GED Instructor | 1994-1997

**Unicom, Inc.** | Systems Programmer and Technical Writer | 1983-1997

## Interests

- Data analytics and statistics
- Programming languages
- The state of the world
- Drawing, painting, guitar, motorcycles, baking, gaming