### 2022-01-30 _academics/ps753-glossary/ps753-glossary.Rmd
PS753 glossary stuff

---
title: "PS753 Glossary"
description: |
  Glossary page for PS753 coursework
author:
  - name: Steve Linberg
    url: {}
date: 2022-01-29
output:
 distill::distill_article:
   self_contained: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

<p style="background: #eee">
[PS 753](../PS753/index.html)
&gt; Glossary
</p>

# Week 1

**From lecture 1a: [Nodes, Edges and Network Samples](https://www.loom.com/share/efc5252bb75b463481b93e55be6e4d67)**

node / vertex
: a junction in a network where two or more lines (edges) intersect; a dot connecting lines.<br>

edge / link
: lines connecting nodes, which indicate some sort of connection or relationship.

node or ego
: $i$ - the node being discussed or focused on

alter
: $j$ - the node that $i$ connects to

network population
: $n$ - the size of the population or count of nodes

With set notation, we define $i$ as a set of $n$ elements:
$$i \in 1, 2, 3 \dots n$$

and $j$ similarly as a set of $n$ elements, except that $j$ cannot equal $i (a node cannot connect to itself):
$$j \in 1, 2, 3 \dots n, i \ne j$$

interaction threshold
: a measure to determine whether two entities have a sufficient connection to be considered to have a link between them

snowball sample
: an entity group formed by starting with one node and "rolling outwards" to its connections until all nodes or actors (in a limited set) are located

egocentric name generator
: a mode of building an entity group (an egocentric network) defined by connection to a single central node; like a snowball sample that doesn't expand past the first set of connections

**From lecture 1b: [Ties and Adjacency Matrices](https://www.loom.com/share/a64c3a975e7147b9bb06bf921325bd96)**

adjacency matrix
: an $n \times n$ matrix depicting connections between $n$ nodes as 0 or 1, where 1 is a connection (or vertex/edge) and 0 is the absence of one:

||A|B|C|D|
|---|---|---|---|---|
|A|-|0|1|0|
|B|0|-|1|0|
|C|1|1|-|0|
|D|0|0|0|-|

directed tie
: a relationship where $W_{i,j} \ne W_{j,i}$, as might be used to represent a transfer of resources from one node to another; graphed with arrows

symmetric tie
: a relationship where $W_{i,j} = W_{j,i}$, with no direction; graphed with lines

binary tie
: a tie where $W_{i,j}$ is 0 or 1, as in the example above, indicating the absence or presence of a tie (also *dichotomous*)

valued tie
: a tie where $W_{i,j}$ is a value $v$, indicating a weight or magnitude of the connection; may be graphed with line attributes such as weight, color, etc

Ties are symmetric or directed, and binary or valued.

**From lecture 1c: [Edgelists](https://www.loom.com/share/9232d0ffe78d4ea4affc543bf4e353af)**

edgelist
: a table indicating edges in a network, with at least "from" and "to" columns, and possibly additional columns for attributes or values

|From|To|Value|
|---|---|---
|A|B|5|
|A|E|2|
|B|A|-1|
|B|C|4|
|B|D|2|
|C|A|-4|

Edgelists are more efficient in sparse networks, as they only list actual connections rather than being a matrix of all possible connections