### 2022-01-30 _academics/ps753/ps753.Rmd
lowercase ps743 you idiot

---
title: "PS753"
description: |
  Political and Social Networks, Spring 2022.
# author:
#   - name: Steve Linberg
# date: 2022-01-29
output:
  distill::distill_article:
    self_contained: false
citation: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

<p style="background: #eee">
PS 753
</p>

Class work from this class will be listed here as we go. This part has to be done by hand in Distill, unfortunately.

- My ongoing [glossary](../ps753-glossary/index.html) page; feel free to comment, or make additions or corrections via comments or github.