### 2022-01-30 _academics/ps753-glossary/ps753-glossary.Rmd
text updates

---
title: "PS753 Glossary"
description: |
  Glossary for PS753 coursework
author:
  - name: Steve Linberg
    url: {}
date: 2022-01-29
output:
 distill::distill_article:
   self_contained: false
   toc: true
bibliography: PS753.bib
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

<p style="background: #eee">
Academics &gt; [PS 753](../ps753/index.html)
&gt; Glossary
</p>

> **A social network** is a set of actors (or points, or nodes, or agents) that may have relationships (or edges, or ties) with one another. Networks can have few or many actors, and one or more kinds of relations between pairs of actors.
>
>-- <cite>[@hanneman, chapter 2]</cite>

## Week 1

### Class lecture 1a: [Nodes, Edges and Network Samples](https://www.loom.com/share/efc5252bb75b463481b93e55be6e4d67)

(Note: some terms overlap somewhat or have context-dependent synonyms.)

node / vertex
: a junction in a network where two or more lines (edges) intersect; a dot connecting lines.

tie / edge / link / relation
: lines connecting nodes, which indicate some sort of connection or relationship.

node / ego / actor
: $i$ - the node being discussed or focused on; @hanneman also note "Actors are described by their relations, not by their attributes."

alter
: $j$ - the node that $i$ connects to

network population
: $n$ - the size of the population or count of nodes

With set notation, we define $i$ as a set of $n$ elements:
$$i \in 1, 2, 3 \dots n$$

and $j$ similarly as a set of $n$ elements, except that $j$ cannot equal $i$ (a node cannot connect to itself):
$$j \in 1, 2, 3 \dots n, i \ne j$$

interaction threshold
: a measure to determine whether two entities have a sufficient connection to be considered to have a link between them

snowball sample
: an entity group formed by starting with "a focal actor or set of actors"^[@hanneman] and "rolling outwards" to its connections until all nodes or actors (in a limited set) are located, or a decision to stop is made

egocentric name generator
: a mode of building an entity group (an egocentric network) defined by connection to a single central node; like a snowball sample that doesn't expand past the first set of connections

### Class lecture 1b: [Ties and Adjacency Matrices](https://www.loom.com/share/a64c3a975e7147b9bb06bf921325bd96)

adjacency matrix
: an $n \times n$ matrix depicting connections between $n$ nodes as 0 or 1, where 1 is a connection (or vertex/edge) and 0 is the absence of one:

||A|B|C|D|
|---|---|---|---|---|
|A|-|0|1|0|
|B|0|-|1|0|
|C|1|1|-|0|
|D|0|0|0|-|

directed tie
: a relationship where $W_{i,j} \ne W_{j,i}$, as might be used to represent a transfer of resources from one node to another; graphed with arrows

symmetric tie
: a relationship where $W_{i,j} = W_{j,i}$, with no direction; graphed with lines

binary tie
: a tie where $W_{i,j}$ is 0 or 1, as in the example above, indicating the absence or presence of a tie (also *dichotomous*)

valued tie
: a tie where $W_{i,j}$ is a value $v$, indicating a weight or magnitude of the connection; may be graphed with line attributes such as weight, color, etc

Ties are symmetric or directed, and binary or valued.

### Class lecture 1c: [Edgelists](https://www.loom.com/share/9232d0ffe78d4ea4affc543bf4e353af)

edgelist
: a table indicating edges in a network, with at least "from" and "to" columns, and possibly additional columns for attributes or values

|From|To|Value|
|---|---|---
|A|B|5|
|A|E|2|
|B|A|-1|
|B|C|4|
|B|D|2|
|C|A|-4|

Edgelists are more efficient in sparse networks, as they only list actual connections rather than being a matrix of all possible connections

### @hanneman, chapter 1 (Social network data):

binary measures of relations
: undirected relations, 0 or 1 for the absence or presence of a connection

multiple-category nominal measures of relations
: directed relations with categories (like "friend, lover, business relationship, kin, or no relationship")

grouped ordinal measures of relations
: measures that reflect a level of intensity or degree; often turned into binary measures by means of a threshold or cut-off

full-rank ordinal measures of relations
: an ordering from 1 to $n$ of an actor's relations (uncommon in social networks)

interval measures of relations
: continuous measures that express the strength of connections by comparison with others, to be able to say "this tie is twice as strong as that tie"; the "most advanced" level of measurement