### 2022-01-31 _academics/ms797-glossary/ms797-glossary.Rmd
Add ms797 index and glossary

---
title: "MS797 Glossary"
description: |
  Glossary for MS797 coursework
author:
  - name: Steve Linberg
    url: {}
date: 2022-01-30
output:
 distill::distill_article:
   self_contained: false
   toc: true
# bibliography: PS753.bib
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

<p style="background: #eee">
Academics &gt; [MS 797](../ms797/index.html)
&gt; Glossary
</p>

Page numbers in parenthesis after terms, from [ISLR](https://www.statlearning.com/) 2nd edition.

## Chapter 2

input variable (15)
: also *predictor*, *independent variable*, *feature*; usually written $X_1, X_2$, etc. The parameter or parameters we are testing to see if they are related to or affect the output.

output variable (15)
: also *response*, *dependent variable*; usually written $Y$. The outcome being measured.

error term (16)
: $\epsilon$ in the equation  
$$Y = f(X) + \epsilon$$
a random quantity of inaccuracy, *independent of X* and with *mean 0*.

systematic (16)
: $f$ in the equation  
$$Y = f(X) + \epsilon$$
the function that describes the (systematic) information $X$ provides about $Y$. This plus the error term equals $Y$.

reducible error (18)
: The amount of the error $\epsilon$ that could be eliminated by improving our estimator $\hat{f}$; the difference between $\hat{f}$ and $f$. This book and course is mostly about ways to minimize the reducible error.

irreducible error (18)
: The amount of $\epsilon$ that could not be reduced even if $f$ was a perfect estimator of $Y$. Always greater than 0. Could be due to hidden variables in $\epsilon$, or random fluctuations in Y, like a measure of "[a] patient's general feeling of well-being on that day".

expected value (19)
: *average value* of an expected measure.

training data (21)
: data used to develop the model for estimating $f$.

parametric methods (21)
: A model based on one or more input parameters, that yields a value for Y, as in:
$$f(X) = \beta_0 + \beta_1X_1 + \beta_2X_2 + \dots + \beta_pX_p$$
$$Y \approx \beta_0 + \beta_1X_1 + \beta_2X_2 + \dots + \beta_pX_p$$
$$\text{income} \approx \beta_0 + \beta_1 \times \text{education} + \beta_2 \times \text{seniority}$$
This creates a predictive, *inflexible* model which usually does not match the true $f$, but which has advantages of simplicity and interpretability. It can be used to predict values for $Y$ based on its parameters, or inputs. Linear and logistic regression are parametric.

non-parametric methods (23)
: methods that do not attempt to estimate $f$. More flexible and have the potential to very closely match observations, but with the risk of *overfitting* the data and increasing the *variance* of subsequent observations. They require much more data than parametric models, and may be difficult to interpret, K-Nearest Neighbor and Support Vector Machines are non-parametric.

prediction (26)
: seeking to guess the value of an response variable $y_i$ given a set of observations and a predictor $f$.

inference (26)
: a model that seems to better understand the relationship between the response and the predictors.

supervised learning (26)
: a category of model that allows us to guess a $y_i$ response to a set of predictor measurements $x_i, i = 1, \dots, n$.

unsupervised learning (26)
: a category of model in which there are observations/measurements $x_i, i = 1, \dots, n$, but no associated response $y_i$. Linear regression cannot be used because there is no response variable to predict.

cluster analysis (27)
: in unsupervised learning, a statistical method for determining whether a set of observations can be divided into "relatively distinct groups," looking for similarities within the groups. (Topic modeling may be an example of this.)

quantitative variables (28)
: numeric values; age, height, weight, quantity. Usually the response variable type for regression problems.

qualitative variables (28)
: also *categorical*: values from a discrete set. Eye color, name, yes/no. Usually the response variable type for classification problems.

regression problems (28)
: problems with quantitative response variables. Given predictors foo, bar, and baz, how big is the frob?

classification problems (28)
: problems with qualitative response variables. Given predictors foo, bar, and baz, is the outcome likely to be a frob, a frib or a freeb?

mean squared error (MSE) (29)
: the average squared error for a set of observations:
$$MSE = \frac{1}{n}\sum_{i=1}^{n}(y_i - \hat{f}(x_i))^2$$
MSE is small if the predicted responses are close to the true responses, and larger as it becomes less accurate; computed from training data, and Gareth *et al.* suggest it should be called *training MSE*.

variance (34)
: _"the amount by which $\hat{f}$ would change if we estimated it using a different training data set"_

bias (35)
: _"the error that is introduced by approximating a real-life problem, which may be extremely complicated, by a much simpler model"_, as in the error from the (presumed) linearity of a regression against non-linear data whose complexity it does not capture. More flexible models increase variance and decrease bias.

bias-variance trade-off (36)
: The tension in seeking the best model for the data between missing the true $f$ with an overly simple (biased) model, vs. an overfitted model with too much variance from mapping too closely to test data.

error rate (37)
: In classification, the proportion of classifications that are mistakes.
$$\frac{1}{n}\sum_{i=1}^{n}I(y_i \neq \hat{y}_i)$$
$I$ is 1 if $y_i \neq \hat{y}_i$ - if the guess for any given $y$ is wrong. The error rate is the percentage of incorrect classifications. Also the *training erorr rate*.

indicator variable (37)
: $I$ in the error rate definition above; a logical variable indicating the presence or absence of a characteristic or trait (such as an accurate classification).

test error rate (37)
: like the *training error rate* but applied to the test data. Uses $\text{Ave}$ instead of sum notation:
$$\text{Ave}(I(y_0 \neq \hat{y}_0))$$
$\hat{y}_0$ is the predicted class label from the classifier.

conditional probability (37)
: The chance that $Y = j$ given an observed $x_0$, as in the *Bayes classifier*:
$$\text{Pr}(Y = j|X = x_0)$$
In a two-class, yes/no classifier, we decide based on whether $\text{Pr}(Y = j|X = x_0)$ is $> 0.5$, or not. Note that $Y$ is the class, as in "ham"/"spam", not a $y$-axis coordinate.

Bayes decision boundary (38)
: a visual depiction of the line of 50% probability dividing (exactly two?) classes in a two-dimensional space

Bayes error rate (38)
: the expected (average) probability of classification error over all values of X in a data set.
$$1 - E(\underset{j}{maxPr}(Y = j|X))$$
The $\underset{j}{maxPr}$ whichever of the $j$ classes has the highest probability for any given value of $X$. Again, $Y$ is not a y-axis coordinate of a two-dimensional space, it's the *class* of the classification: "yes"/"no", "ham"/"spam", "infected"/"not infected". Also:
_"The Bayes error rate is analogous to the irreducible error, discussed earlier."_

K-nearest-neighbors (KNN) (39)
: a classifier that assigns a class Y to an observation based on the population proportions of its nearest neighbors; a circular "neighborhood" on a two-dimensional plot. It looks at actual data points that have been classified, and asks what any given non-classified point would be classified as based on its nearest neighbors.

## Chapter 3

Synergy effect / interaction effect (60)
: when two or more predictors affect each other as well as the outcome; when 50k each in TV or radio ads give different results than 100k in either one

Simple linear regression
: the simplest model, predicting $Y$ from a single predictor $X$.
$$Y \approx \beta_0 + \beta_1X$$
$\approx$ = "is approximately modeled as"

least squares (61)
: the most common measure of closeness of a regression line to its data points, the sum of squares of the distances between the points and the closest point on the line (directly above or below)

residual (61)
: the difference between $y_i$ and $\hat{y}_i$, also $e_i$; the difference between the $i$th response variable and the $i$th response variable predicted by the model

residual sum of squares (RSS) (62)
: the sum of the squared residuals for each point on the regression line
$$\text{RSS} = e_1^2 + e_2^2 + \dots + e_n^2$$
Formulas for $\hat{\beta_0}$ and $\hat{\beta_1}$ are on p. 62

intercept ($\beta_0$) (63)
: the expected value of $Y$ when $X = 0$

slope ($\beta_1$) (63)
: the average increase in $Y$ associated with a one-unit increase in $X$

error term ($\epsilon$) (63)
: whatever we missed with the model, due to the true model not being linear (it almost never is), measurement error, or other variables that cause variation in $Y$

population regression line (63)
: "the best linear approximation to the true relationship between $X$ and $Y$"
$$Y = \beta_0 + \beta_1X + \epsilon$$
least squares line (63)
: the regression line made of the least-squares estimates for $\beta_0$ and $\beta_1$

bias (65)
: in an estimator, something that systematically misses the true parameter; for an unbiased estimator, $\hat{\mu} = \mu$ when averaged over (huge) numbers of observations

standard error (SE) (65)
: the average amount that an estimate $\hat{\mu}$ (sample mean) differs from the actual value of $\mu$ (population mean)
$$\text{Var}(\hat{\mu}) = \text{SE}(\hat{\mu})^2 = \frac{\sigma^2}{n} (\text{also} = \frac{\sigma}{\sqrt{n}})$$
$\sigma$ is "the standard deviation of each of the realizations $y_i$ of $Y$. Since $\sigma^2$ is divided by $n$, the standard error shrinks as observations increase. It represents the amount we would expect means of additional samples to "jump around" simply due to random chance and the limitations of the model's accuracy.^[http://faculty.ucr.edu/~hanneman/nettext/C1_Social_Network_Data.html]

residual standard error (RSE) (66)
: the estimate of $\sigma$
$$\text{RSE} = \sqrt{RSS / (n-2)}$$

confidence interval (66)
: a range of values within which we have a measured probability (often 95%) of containing the true value of the parameter; a 95% confidence interval in linear regression takes the form
$$\hat{\beta_1} \pm 2 \cdot \text{SE}(\hat{\beta_1})$$

t-statistic (67)
: the number of standard deviations that $\hat{\beta_1}$ is away from $0$.
$$t = \frac{{\hat{\beta_1}} - 0}{\text{SE}(\hat{\beta_1})}$$
For there to be a relationship between $X$ and $Y$, $\hat{\beta_1}$ has to be nonzero (i.e. have a slope). The standard error (SE) of $\hat{\beta_1}$ (in the denominator above) measures its accuracy; if it is small, then $t$ will be larger, and if it is large, then $t$ will be smaller. $t$ is around 2 for a p-value of 0.05 (actually about 1.96, as 2 standard deviations is 95.45% of a normal distribution), and around 2.75 for a p-value of 0.01.

p-value (67)
: the probability of observing a value greater than $|t|$ by chance.

total sum of squares (TSS) (70)
: the total variance in the response $Y$
$$\text{TSS} = \sum(y_i - \bar{y})^2$$
compare with RSS, the amount of variability left unexplained after the regression. TSS - RSS is the amount of variability (or error) explained by the regression.

$R^2$ statistic (70)
: the proportion of variance in $Y$ explained by $X$, a range from 0 to 1
$$R^2 = \frac{\text{TSS - RSS}}{\text{TSS}} = 1 - \frac{\text{RSS}}{\text{TSS}}$$
$R^2$ values close to 1 indicate a regression that explains a lot of the variability in the response, and a stronger model. A value close to 0 indicates that the regression doesn't explain much of the variability.

correlation (70)
: a measure of the linearity of the relationship between $X$ and $Y$; values close to 0 indicate weak-to-no relationship, values near 1 or -1 indicate strong positive or negative correlation
(horrible formula on p.70)