### 2022-01-31 _posts/2022-01-30-ms797-glossary-underway/ms797-glossary-underway.Rmd
New post

---
title: "MS797 Glossary underway"
description: |
  The other big glossary for this semester is underway.
author:
  - name: Steve Linberg
    url: https://slinberg.net
date: 2022-01-30
output:
  distill::distill_article:
    self_contained: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

The term's other major [glossary](../../academics/ms797-glossary/), for [MS797, Machine Learning for Social Science](../../academics/ms797), is up. At this point it's about halfway through chapter 3... lots to go. This was initially in a standalone repo, but I think it makes more sense to migrate it here.

Again, comments, clarifications or corrections are all welcome.