### 2022-02-01 _academics/ps753-glossary/ps753-glossary.Rmd
Add week 2 terms

---
title: "PS753 Glossary"
description: |
  Glossary for PS753 coursework
author:
  - name: Steve Linberg
    url: {}
date: 2022-01-29
output:
 distill::distill_article:
   self_contained: false
   toc: true
bibliography: PS753.bib
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

<p style="background: #eee">
Academics &gt; [PS 753](../ps753/index.html)
&gt; Glossary
</p>

> **A social network** is a set of actors (or points, or nodes, or agents) that may have relationships (or edges, or ties) with one another. Networks can have few or many actors, and one or more kinds of relations between pairs of actors.
>
>-- <cite>[@hanneman, chapter 2]</cite>

## Week 1

### Class lecture 1a: [Nodes, Edges and Network Samples](https://www.loom.com/share/efc5252bb75b463481b93e55be6e4d67)

(Note: some terms overlap somewhat or have context-dependent synonyms.)

node / vertex
: a junction in a network where two or more lines (edges) intersect; a dot connecting lines.

tie / edge / link / relation
: lines connecting nodes, which indicate some sort of connection or relationship.

node / ego / actor
: $i$ - the node being discussed or focused on; also note "Actors are described by their relations, not by their attributes."^[@hanneman]

alter
: $j$ - the node that $i$ connects to

network population
: $n$ - the size of the population or count of nodes

With set notation, we define $i$ as a set of $n$ elements:
$$i \in 1, 2, 3 \dots n$$

and $j$ similarly as a set of $n$ elements, except that $j$ cannot equal $i$ (a node cannot connect to itself):
$$j \in 1, 2, 3 \dots n, i \ne j$$

interaction threshold
: a measure to determine whether two entities have a sufficient connection to be considered to have a link between them

snowball sample
: an entity group formed by starting with "a focal actor or set of actors"^[@hanneman] and "rolling outwards" to its connections until all nodes or actors (in a limited set) are located, or a decision to stop is made

egocentric name generator
: a mode of building an entity group (an egocentric network) defined by connection to a single central node; like a snowball sample that doesn't expand past the first set of connections

### Class lecture 1b: [Ties and Adjacency Matrices](https://www.loom.com/share/a64c3a975e7147b9bb06bf921325bd96)

adjacency matrix
: an $n \times n$ matrix depicting connections between $n$ nodes as 0 or 1, where 1 is a connection (or vertex/edge) and 0 is the absence of one:

||A|B|C|D|
|---|---|---|---|---|
|A|-|0|1|0|
|B|0|-|1|0|
|C|1|1|-|0|
|D|0|0|0|-|

directed tie
: a relationship where $W_{i,j} \ne W_{j,i}$, as might be used to represent a transfer of resources from one node to another; graphed with arrows

symmetric tie
: a relationship where $W_{i,j} = W_{j,i}$, with no direction; graphed with lines

binary tie
: a tie where $W_{i,j}$ is 0 or 1, as in the example above, indicating the absence or presence of a tie (also *dichotomous*)

valued tie
: a tie where $W_{i,j}$ is a value $v$, indicating a weight or magnitude of the connection; may be graphed with line attributes such as weight, color, etc

Ties are symmetric or directed, and binary or valued.

### Class lecture 1c: [Edgelists](https://www.loom.com/share/9232d0ffe78d4ea4affc543bf4e353af)

edgelist
: a table indicating edges in a network, with at least "from" and "to" columns, and possibly additional columns for attributes or values

|From|To|Value|
|---|---|---
|A|B|5|
|A|E|2|
|B|A|-1|
|B|C|4|
|B|D|2|
|C|A|-4|

Edgelists are more efficient in sparse networks, as they only list actual connections rather than being a matrix of all possible connections

### @hanneman, chapter 1 (Social network data):

binary measures of relations
: undirected relations, 0 or 1 for the absence or presence of a connection

multiple-category nominal measures of relations
: directed relations with categories (like "friend, lover, business relationship, kin, or no relationship")

grouped ordinal measures of relations
: measures that reflect a level of intensity or degree; often turned into binary measures by means of a threshold or cut-off

full-rank ordinal measures of relations
: an ordering from 1 to $n$ of an actor's relations (uncommon in social networks)

interval measures of relations
: continuous measures that express the strength of connections by comparison with others, to be able to say "this tie is twice as strong as that tie"; the "most advanced" level of measurement

### @lazer2011

homophily
: the idea that individuals who are similar to one another are more likely to form ties

whole network data
: relational information about a whole set (or subset) of data, with all of the actors' relations to each other considered

egocentric data
: relational information about a set of nodes connected to one particular node and not to each other

diameter
: the maximum degree of separation between any two nodes in the network

one-mode data
: ties among one set (or category) of agent, such as nations in the context of trade

two-mode data
: ties between different sets (or categories) of agents, such as ties between nations and international organizations

### @borgatti2009

sociometry
: "a technique for eliciting and graphically representing individuals' subjective feelings toward one another"

strength of weak ties (SWT) theory
: the theory that one is likelier to hear new information from people they aren't closely connected to in a network (c.f. homiphily)

centrality
: a family of positional properties of nodes in a network

Freeman's betweenness
: a type of centrality where a node is frequently along the shortest path between pairs of nodes, giving control over flow or power

opportunity-based antecedents
: "the likelihood that two nodes will come into contact" - when considering the formation of ties

benefit-based antecedents
: "some kind of utility maximization or discomfort minimization that leads to tie formation"

node homogeneity
: a category of node outcomes referring to the similarity of nodes

performance
: a category of node outcomes referring to some good (such as strong performance)

## Week 2

### Class lecture 2a: [Network Structure - Walks, Paths, and Distance](https://www.loom.com/share/a8a7949244f04ed5995d572340c20aa0)

Connections between nodes:

adjacent
: a direct connection between nodes; not necessarily bilateral in directed networks. A leading to B ("A adj B") does not mean that B leads to A ("B adj A)

reachable
: whether a node is reachable from another node, regardless of distance

distance
: the number of ties (steps, edges) that must be traversed to reach a target node

walk
: sequence (not *path*, see below) that connects two nodes, consisting of the nodes and edges

trail
: a walk that can only go through each edge / tie once, but can hit the same node more than once

path
: a trail that only hits each node and edge once; the start and end node may be the same

geodesic distance
: the shortest path between two nodes; by definition, not a trail or walk because repeated segments wouldn't be the shortest; with binary data, the number of steps; with weighted data, might be a sum or some other calculation of "effort"

### Class lecture 2b: [Graph Substructures and Components](https://www.loom.com/share/515068ae6dc641be984c0ef974fb8e52)

Network substructures

dyad, triad, clique
: two, three, or four-or-more connected nodes

complete graph
: a network where every node is directly connected to every other node

connected graph
: a network where every node is indirectly connected to every other node

unconnected graph
: a network where at least one node is unreachable

component
: the set of all points that constitutes a connected subgraph within a network

main component
: the largest component within a network

minor component
: a smaller one, possibly one of many

pendant
: a node with only one link or edge to a network, "dangling"

isolate
: an unconnected node

### Class lecture 2c: [Dyad and Triad Census](https://www.loom.com/share/96aaf1d2982d427c83564c338d2e0b83)

mutual dyad
: a dyad where both nodes connect to each other, as in an undirected network

asymmetric dyad
: a dyad (in a directed network) where one node connects to another, but non-reciprocally (one way only)

null
: a dyad of two unconnected nodes

(empty / one edge / two path / triangle) triad
: a triad with zero, one, two or three edges between three nodes (all four possible permutations)

balance theory
: the theory that two nodes connected to a common node will also develop connections to each other

global transitivity index
: the proportion of triads in a network that are complete (with 3 connections between them)

There is a vocabulary for triads in directed networks describing the 16 possible permutations of ties (or the absence thereof) among 3 nodes, counting the number of mutual, asymmetric and null dyads, with direction indicators, like `003` or `120D` - see the slide at 3:45

vacuously transitive triad
: a triad where (5 of 16 possibilities)

intransitive triad
: a triad where (7 of 16 possibilities)

transitive triad
: a triad where (4 of 16 possibilities)

xxx see https://corescholar.libraries.wright.edu/cgi/viewcontent.cgi?article=2427&context=knoesis

### Class lecture 2d: [Transitivity and Clustering Coefficient](https://www.loom.com/share/ff29331e900741dd8634d921e4abafdc)

local transitivity / local clustering coefficient
: the likelihood that the neighbors of a node are also connected to each other; the number of connections that *do* exist over the number of connections that *could* exist