### 2022-02-02 _posts/2022-02-02-pythagorean-theorem/pythagorean-theorem.Rmd
new post

---
title: "Pythagorean Theorem"
description: |
  How did I not know this
author:
  - name: Steve Linberg
    url: https://slinberg.net
date: 2022-02-02
output:
  distill::distill_article:
    self_contained: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

How did I make it to `$adulthood * 2` without knowing that our old friend, the Pythagorean Theorem ($a^2 + b^2 = c^2$), works in any number of dimensions?

The Euclidean distance $d$ between the origin ($0, 0, \dots 0$) is just

$$d = \sqrt{x_1^2 + x_2^2 + \dots + x_n^2}$$

Add up all of the squares and take the square root. Dope-smack for me.