### 2022-02-07 _academics/ms797-useful-r-tidbits/ms797-useful-r-tidbits.Rmd
build

---
title: "Useful R tidbits"
description: |
  Tasty R snacks that do useful things
author:
  - name: Steve Linberg
    url: {}
date: 2022-02-06
output:
  distill::distill_article:
    self_contained: false
    toc: true
preview: ISLR-cover.png
---

```{r setup, include=FALSE}
# knitr::opts_chunk$set(echo = FALSE)
```

<p style="background: #eee">
Academics &gt; [MS797](../ms797/index.html)
&gt; Useful R tidbits
</p>

Most of this comes directly from the [resources from the 2nd edition](https://www.statlearning.com/resources-second-edition) of [ISLR](https://www.statlearning.com/), either direct quotes or things that build on them. The `Lab` section of each chapter contains useful walkthroughs and code illustrating the chapter's key concepts, and there's always something to learn about R and how different people use it.

<aside>
[![ISLRv2](ISLR-cover.png){width=150px}](https://www.statlearning.com)
</aside>

Everything here is, unless otherwise marked, either the work of, or derivative of the work of, the original authors of the book.

## Chapter 3: Linear Regression

```{r chunk1, echo=FALSE}
library(MASS)
library(ISLR2)
```
### names(): see the attributes of a model

`names()` is a nice, simple way to see the attributes of a linear model:

```{r chunk4}
attach(Boston)
lm.fit <- lm(medv ~ lstat)
names(lm.fit)
```

### coef(): see the model's coefficients

`coef()` is a convenient way to get the `coefficients` attribute:

```{r chunk6}
coef(lm.fit)
```

Note: this is equivalent to `lm.fit$coefficients`, but slightly easier on the eyes.

### confint(): see confidence intervals for a model

```{r chunk7}
confint(lm.fit)
```

Note: levels can be added with the `level` parameter (the default is 0.95):

```{r}
confint(lm.fit, level = 0.90)
```

### predict(): see confidence OR prediction intervals for a model

```{r chunk8}
predict(lm.fit, data.frame(lstat = (c(5, 10, 15))),
    interval = "confidence")
predict(lm.fit, data.frame(lstat = (c(5, 10, 15))),
    interval = "prediction")
```

Note: prediction intervals are defined on p. 82; in brief, it is the interval in which we are $X$% certain that any future observation will fall.

Note: the `data.frame()` call in the above parameter lists produce a nifty little data frame that is used by `predict` to identify the column to do calculations on, and the desired levels:

```{r}
data.frame(lstat = (c(5, 10, 15)))
```

Note: the book uses the base R graphics:

```{r chunk9}
plot(lstat, medv)
abline(lm.fit)
```

The tidyverse/ggplot version of this is:

```{r chunk9a}
library(tidyverse)
ggplot(Boston, mapping = aes(x = lstat, y = medv)) +
  geom_point() +
  geom_abline(intercept = coef(lm.fit)[1],
              slope = coef(lm.fit)[2])
```

...a bit more work, but more modern.

### arranging plots in grids

The book uses base R, again, with `par()` and `mfrow()` to arrange plots:

```{r chunk11}
par(mfrow = c(2, 2))
plot(lm.fit)
```

ggplot uses `facet()` to arrange plots from the same data set in grids.

Plots from different data sets need additional packages to be combined; one option is the `cowplot` library and the `plot_grid()` function.

XXX
<hr>

Alternatively, we can compute the residuals from a linear regression
fit using the `residuals()` function. The function
`rstudent()` will return the studentized residuals, and we
can use this function to plot the residuals against the fitted values.

```{r chunk12}
plot(predict(lm.fit), residuals(lm.fit))
plot(predict(lm.fit), rstudent(lm.fit))
```

On the basis of the residual plots, there is some evidence of non-linearity.
Leverage statistics can be computed for any number of predictors using the `hatvalues()` function.

```{r chunk13}
plot(hatvalues(lm.fit))
which.max(hatvalues(lm.fit))
```

The `which.max()` function identifies the index of the largest element of a vector. In this case, it tells us which observation has the largest leverage statistic.

## Multiple Linear Regression

In order to fit a multiple linear regression model using least squares, we again use the `lm()` function. The syntax {\R{lm(y $\sim$ x1 + x2 + x3)}} is used to fit a model with three predictors, `x1`, `x2`, and `x3`.
The `summary()` function now outputs the regression coefficients for all the predictors.

```{r chunk14}
lm.fit <- lm(medv ~ lstat + age, data = Boston)
summary(lm.fit)
```

The `Boston` data set contains 12 variables, and so it would be cumbersome to have to type all of these in order to perform a regression using all of the predictors.
Instead, we can use the following short-hand:

```{r chunk15}
lm.fit <- lm(medv ~ ., data = Boston)
summary(lm.fit)
```

We can access the individual components of a summary object by name
(type `?summary.lm` to see what is available). Hence
`summary(lm.fit)$r.sq` gives us the $R^2$, and
`summary(lm.fit)$sigma` gives us the RSE. The `vif()`
function, part of the `car` package, can be used to compute variance
inflation factors.   Most VIF's are
low to moderate for this data. The `car` package is not part of the base `R` installation so it must be downloaded the first time you use it via the `install.packages()` function in `R`.

```{r chunk16}
library(car)
vif(lm.fit)
```

What if we would like to perform a regression using all of the variables but one?  For example, in the above regression output,  `age` has a high $p$-value. So we may wish to run a regression excluding this predictor.
 The following syntax results in a regression using all predictors except `age`.

```{r chunk17}
lm.fit1 <- lm(medv ~ . - age, data = Boston)
summary(lm.fit1)
```

Alternatively, the `update()` function can be used.

```{r chunk18}
lm.fit1 <- update(lm.fit, ~ . - age)
```

## Interaction Terms

It is easy to include interaction terms in a linear model using the `lm()` function. The syntax `lstat:black` tells `R` to include an interaction term between `lstat` and `black`.
The syntax `lstat * age` simultaneously includes `lstat`, `age`, and the interaction term `lstat`$\times$`age` as predictors; it is a shorthand for `lstat + age + lstat:age`.
  %We can also pass in transformed versions of the predictors.

```{r chunk19}
summary(lm(medv ~ lstat * age, data = Boston))
```

## Non-linear Transformations of the Predictors

The `lm()` function can also accommodate non-linear transformations of the predictors. For instance, given a predictor $X$, we can create a predictor $X^2$ using
 `I(X^2)`. The function `I()` is needed since the `^` has a special meaning in a formula object; wrapping as we do allows the standard usage in `R`, which is to raise `X` to the power `2`. We now
perform a regression of `medv` onto `lstat` and `lstat^2`.

```{r chunk20}
lm.fit2 <- lm(medv ~ lstat + I(lstat^2))
summary(lm.fit2)
```

The near-zero $p$-value associated with the quadratic term suggests that it leads to an improved model.
We use the `anova()` function  to further quantify the extent to which the quadratic fit is superior to the linear fit.

```{r chunk21}
lm.fit <- lm(medv ~ lstat)
anova(lm.fit, lm.fit2)
```

Here Model 1 represents the linear submodel containing only one predictor, `lstat`, while Model 2 corresponds to the larger quadratic model that has two predictors, `lstat` and `lstat^2`.
The `anova()` function performs a hypothesis test
comparing the two models. The   null hypothesis is that the two models fit the data equally well,  and the alternative hypothesis is that the full model is superior. Here the $F$-statistic is $135$
 and the associated $p$-value is virtually zero. This provides very clear evidence that the model containing the predictors `lstat` and `lstat^2` is far superior to the model that only contains the predictor `lstat`.
 This is not surprising, since earlier we saw evidence for non-linearity in the relationship between `medv` and `lstat`. If we type

```{r chunk22}
par(mfrow = c(2, 2))
plot(lm.fit2)
```

 then we see that when the `lstat^2` term is included in the model, there is little discernible pattern in the residuals.

In order to create a cubic fit, we can include a predictor of the form `I(X^3)`. However, this approach can start to get cumbersome for higher-order polynomials. A better approach involves using the `poly()` function to create the polynomial within `lm()`. For example, the following command produces a
fifth-order polynomial fit:

```{r chunk23}
lm.fit5 <- lm(medv ~ poly(lstat, 5))
summary(lm.fit5)
```

This suggests that including additional  polynomial terms, up to fifth order, leads to an improvement in the model fit! However, further investigation of the data reveals that no polynomial terms beyond fifth order have significant $p$-values
in a regression fit.

 By default, the `poly()` function orthogonalizes the predictors:
 this means that the features output by this function are not simply a
 sequence of powers of the argument. However, a linear model applied to the output of the `poly()` function will have the same fitted values as a linear model applied to the raw polynomials (although the coefficient estimates, standard errors, and p-values will differ). In order to obtain the raw polynomials from the `poly()` function,  the argument `raw = TRUE` must be used.

Of course, we are in no way restricted to using polynomial transformations of the predictors. Here we try a log transformation.

```{r chunk24}
summary(lm(medv ~ log(rm), data = Boston))
```

## Qualitative Predictors

We will now examine the `Carseats` data, which is part of the
`ISLR2` library. We will  attempt to predict `Sales`
(child car seat sales) in $400$ locations based on a number of
predictors.

```{r chunk25}
head(Carseats)
```

The `Carseats` data includes qualitative predictors such as `shelveloc`, an indicator of the quality of the shelving location---that is, the  space within a store in which the car seat is displayed---at each location. The predictor `shelveloc` takes on three possible values:  *Bad*, *Medium*, and *Good*. Given a qualitative variable such as `shelveloc`, `R` generates dummy variables automatically. Below we fit a multiple regression model that includes some interaction terms.

```{r chunk26}
lm.fit <- lm(Sales ~ . + Income:Advertising + Price:Age, 
    data = Carseats)
summary(lm.fit)
```

The `contrasts()` function returns the coding that `R` uses for the dummy variables.

```{r chunk27}
attach(Carseats)
contrasts(ShelveLoc)
```

Use `?contrasts` to learn about other contrasts, and how to set them.

`R` has created a `ShelveLocGood` dummy variable that takes on a value of 1 if the shelving location is good, and 0 otherwise. It has also created a `ShelveLocMedium` dummy variable that equals 1 if the shelving location is medium, and 0 otherwise.
A bad shelving location corresponds to a zero for each of the two dummy variables.
The fact that the coefficient for `ShelveLocGood` in the regression output is positive indicates that a good shelving location is associated with high sales (relative to a bad location). And `ShelveLocMedium` has a smaller positive coefficient, indicating that a medium shelving location is associated with higher sales than a bad shelving location but lower sales than a good shelving location.