### 2022-02-08 _academics/ms797/ms797.Rmd
build

---
title: "MS797"
description: |
  Machine Learning for Social Sciences, Spring 2022.
# author:
#   - name: Steve Linberg
# date: 2022-01-30
output:
  distill::distill_article:
    self_contained: false
citation: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

<p style="background: #eee">
[<i class="fas fa-home"></i>](/) &gt; Academics &gt; MS797
</p>

Work and assignments from this class will be added here as we go. (This part has to be done by hand in Distill, unfortunately.)

- My ongoing [glossary](../ms797-glossary/index.html) page; feel free to comment, or make additions or corrections via comments or github.
- Some [useful R tidbits](../ms797-useful-r-tidbits/index.html) from, or inspired by, the text.