### 2022-02-08 _academics/ps753-useful-r-tidbits/ps753-useful-r-tidbits.Rmd

---
title: "Useful R tidbits: Social Networks"
description: |
  Tasty R snacks that do useful things in Social Networks
author:
  - name: Steve Linberg
    url: {}
date: 2022-02-08
output:
  distill::distill_article:
    self_contained: false
    toc: true
---

```{r setup, include=FALSE}
# knitr::opts_chunk$set(echo = FALSE)
```

<p style="background: #eee">
[<i class="fas fa-home"></i>](/) &gt; Academics &gt; [PS753](../ps753/index.html)
&gt; Useful R tidbits
</p>

This document contains R snippets pertaining to work in the [Social Networks class](../ps753/index.html). There is also one for the [Machine Learning class](../ms797-useful-r-tidbits/index.html).