### 2022-02-17 _academics/ps753-hw4/ps753-hw4.Rmd
Initial commit

---
title: "Homework 4: Centrality Analysis of the Florentine Wedding dataset"
description: |
  A first-look network analysis
author:
  - name: Steve Linberg
    url: {}
date: 2022-02-09
output:
  distill::distill_article:
    self_contained: false
    code_folding: true
categories:
  - r
  - social networks
---

<p style="background: #eee">
[<i class="fas fa-home"></i>](/) &gt; Academics &gt; [PS753](../ps753/index.html)
&gt; Homework 4: Centrality Analysis of the Florentine Wedding dataset
</p>

```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = FALSE,
	fig.height = 8,
	message = FALSE,
	warning = FALSE
)
library(network)
library(igraph)
library(tidyverse)
```

```{r}
data(flo)
ntwk.ig <- graph_from_adjacency_matrix(flo)

ntwk.vcount <- vcount(ntwk.ig)
ntwk.ecount <- ecount(ntwk.ig)
ntwk.is_bipartite <- is_bipartite(ntwk.ig)
ntwk.is_directed <- is_directed(ntwk.ig)
ntwk.is_weighted <- is_weighted(ntwk.ig)
ntwk.vertex_attr_names <- vertex_attr_names(ntwk.ig)
ntwk.edge_attr_names <- edge_attr_names(ntwk.ig)
ntwk.names <- V(ntwk.ig)$name
ntwk.dyad_census <- igraph::dyad.census(ntwk.ig)
ntwk.triad_census <- igraph::triad_census(ntwk.ig)
ntwk.global_clustering_coef <- transitivity(ntwk.ig, type = "global")
ntwk.local_avg_clustering_coef <- transitivity(ntwk.ig, type = "average")
ntwk.avg_path_length <- average.path.length(ntwk.ig, directed = T)
ntwk.component_names <- names(igraph::components(ntwk.ig))
ntwk.components_no <- igraph::components(ntwk.ig)$no
ntwk.components_csize <- igraph::components(ntwk.ig)$csize

# distances(ntwk.ig,"Acciaiuoli","Strozzi", weights=NA)

ntwk.graph_density <- graph.density(ntwk.ig)
ntwk.graph_density_noloops <- graph.density(ntwk.ig, loops = TRUE)

ntwk.avg_degree <- igraph::degree(ntwk.ig)

# Get the degree info and make the data frame

ntwk.nodes <- data.frame(
  name = V(ntwk.ig)$name, 
  degree = igraph::degree(ntwk.ig),
  indegrees = igraph::degree(ntwk.ig, mode = "in", loops = FALSE),
  outdegrees = igraph::degree(ntwk.ig, mode = "out", loops = FALSE)
  )
# ntwk.nodes[ntwk.nodes$name %in% c("Medici", "Strozzi"),]

# Get degree centralization

ntwk.indegree_centralization <- 
  centr_degree(ntwk.ig, loops = FALSE, mode = "in")$centralization
ntwk.outdegree_centralization <- 
  centr_degree(ntwk.ig, loops = FALSE, mode = "out")$centralization
ntwk.total_centralization <- 
  centr_degree(ntwk.ig, loops = FALSE, mode = "total")$centralization
```

This week, we're going to look at various measures of *degree centrality* in the Florentine Wedding dataset.