### 2022-03-20 _academics/ps753-hw5/ps753-hw5.Rmd
New post

---
title: "Homework 5: Brokerage and Betweenness in the Florentine Wedding dataset"
description: |
  Examining additional centrality measures in the Florentine network
author:
  - name: Steve Linberg
    url: {}
date: 2022-02-16
output:
  distill::distill_article:
    self_contained: false
    code_folding: true
    toc: true
categories:
  - r
  - social networks
---

<p style="background: #eee">
[<i class="fas fa-home"></i>](/) &gt; Academics &gt; [PS753](../ps753/index.html)
&gt; `r rmarkdown::metadata$title`
</p>

```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = FALSE,
	fig.height = 8,
	message = FALSE,
	warning = FALSE
)
library(network)
library(igraph)
library(tidyverse)
```