### 2022-03-20 _academics/ps753/ps753.Rmd
Add to index:

---
title: "PS753"
description: |
  Political and Social Networks, Spring 2022.
# author:
#   - name: Steve Linberg
# date: 2022-01-29
output:
  distill::distill_article:
    self_contained: false
citation: false
categories:
  - social networks
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

<p style="background: #eee">
[<i class="fas fa-home"></i>](/) &gt; Academics &gt; PS 753
</p>

Work and assignments from this class will be added here as we go. (This part has to be done by hand in Distill, unfortunately.)

- My ongoing [glossary](../ps753-glossary/index.html) page; feel free to comment, or make additions or corrections via comments or github.
- Some [useful R tidbits](../ps753-useful-r-tidbits/index.html) from the course material and related sources.
- [Homework 3](../ps753-hw3/index.html) for the course: a network analysis of the Florentine Wedding data.
- [Homework 4](../ps753-hw4/index.html) for the course: a centrality analysis of the Florentine Wedding data.
- [Homework 5](../ps753-hw5/index.html) for the course: brokerage and betweenness analysis of the Florentine Wedding data.