### 2022-03-20 _academics/ps753-hw5/ps753-hw5.Rmd
knit

---
title: "Homework 5: Brokerage and Betweenness in the Florentine Wedding dataset"
description: |
  Examining additional centrality measures in the Florentine network
author:
  - name: Steve Linberg
    url: {}
date: 2022-02-16
output:
  distill::distill_article:
    self_contained: false
    code_folding: true
    toc: true
categories:
  - r
  - social networks
---

<p style="background: #eee">
[<i class="fas fa-home"></i>](/) &gt; Academics &gt; [PS753](../ps753/index.html)
&gt; `r rmarkdown::metadata$title`
</p>

```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = FALSE,
	fig.height = 8,
	message = FALSE,
	warning = FALSE
)
library(network)
library(igraph)
library(tidyverse)
```

```{r}
data(flo)
ntwk.ig <- graph_from_adjacency_matrix(flo)

ntwk.vcount <- vcount(ntwk.ig)
ntwk.ecount <- ecount(ntwk.ig)
ntwk.is_bipartite <- is_bipartite(ntwk.ig)
ntwk.is_directed <- is_directed(ntwk.ig)
ntwk.is_weighted <- is_weighted(ntwk.ig)
ntwk.vertex_attr_names <- vertex_attr_names(ntwk.ig)
ntwk.edge_attr_names <- edge_attr_names(ntwk.ig)
ntwk.names <- V(ntwk.ig)$name
ntwk.dyad_census <- igraph::dyad.census(ntwk.ig)
ntwk.triad_census <- igraph::triad_census(ntwk.ig)
ntwk.global_clustering_coef <- transitivity(ntwk.ig, type = "global")
ntwk.local_avg_clustering_coef <- transitivity(ntwk.ig, type = "average")
ntwk.avg_path_length <- average.path.length(ntwk.ig, directed = T)
ntwk.component_names <- names(igraph::components(ntwk.ig))
ntwk.components_no <- igraph::components(ntwk.ig)$no
ntwk.components_csize <- igraph::components(ntwk.ig)$csize

# distances(ntwk.ig,"Acciaiuoli","Strozzi", weights=NA)

ntwk.graph_density <- graph.density(ntwk.ig)
ntwk.graph_density_noloops <- graph.density(ntwk.ig, loops = TRUE)

ntwk.avg_degree <- igraph::degree(ntwk.ig)

# Get the degree info and make the data frame

ntwk.nodes <- data.frame(
  name = V(ntwk.ig)$name, 
  degree = igraph::degree(ntwk.ig),
  indegrees = igraph::degree(ntwk.ig, mode = "in", loops = FALSE),
  outdegrees = igraph::degree(ntwk.ig, mode = "out", loops = FALSE)
  )
# ntwk.nodes[ntwk.nodes$name %in% c("Medici", "Strozzi"),]

# Get degree centralization

ntwk.indegree_centralization <- 
  centr_degree(ntwk.ig, loops = FALSE, mode = "in")$centralization
ntwk.outdegree_centralization <- 
  centr_degree(ntwk.ig, loops = FALSE, mode = "out")$centralization
ntwk.total_centralization <- 
  centr_degree(ntwk.ig, loops = FALSE, mode = "total")$centralization
```

Building on [last week's look](../ps753-hw3/index.html) at basic network centrality attributes in the Florentine Wedding dataset, we're going to look at two new measurements of centrality: brokerage and betweenness.

One of the under-the-hood consequences of this is that we must switch from iGraph to Statnet for this analysis, as brokerage is only found in Stanet.

```{r}
# statnet version
climinfl.nodes <- data.frame(
    name      = climinfl.stat %v% "vertex.names",
    totdegree = sna::degree(climinfl.stat),
    indegree  = sna::degree(climinfl.stat, cmode = "indegree"),
    outdegree = sna::degree(climinfl.stat, cmode = "outdegree"),
    eigen     = sna::evcent(climinfl.stat, ignore.eval = TRUE),
    bonanich  = sna::bonpow(climinfl.stat),
    centr_clo = sna::closeness(climinfl.stat, gmode = "graph"),
    centr_btw = sna::betweenness(climinfl.stat, gmode = "graph")
)
# Network-level measures:
#   closeness centralization
climinfl.centr_clo = centralization(climinfl.stat, FUN = "closeness", mode = "graph")
#   betweenness centralization
climinfl.centr_btw = centralization(climinfl.stat, FUN = "betweenness", mode = "graph")

# Statnet-only: Gould-Fernandez Brokerage
temp <- data.frame(brokerage(climinfl.stat, cl = climinfl.nodes$orgtype5)$z.nli)
climinfl.nodes <- climinfl.nodes %>%
  mutate(broker.tot = temp$t,
         broker.coord = temp$w_I,
         broker.itin = temp$w_O,
         broker.rep = temp$b_IO,
         broker.gate = temp$b_OI,
         broker.lia = temp$b_O)
```