### 2022-08-18 steve.Rmd
text updates

---
title: "Steve Linberg"
image: "2022-sfl-photo-400x400.jpg"
links:
  - label: LinkedIn
    url: "https://www.linkedin.com/in/slinberg"
  - label: Twitter
    url: "https://twitter.com/stevelinberg"
  - label: GitHub
    url: "https://github.com/stevelinberg"
  - label: SO
    url: "https://stackoverflow.com/users/13603796/stevelinberg"
  - label: Email
    url: "mailto:steve@slinberg.net"
output:
  postcards::trestles
---

## Bio

In summer 2022, I completed the [Data Analytics and Computational Social Science (DACSS) Masters program](https://www.umass.edu/sbs/data-analytics-and-computational-social-science-program) at the [University of Massachusetts, Amherst](https://umass.edu). My present position is Senior Systems and Analytics Engineer at the [Gandara Center](https://www.gandaracenter.org/) in West Springfield, MA.

This site contains my academic [blog](index.html) with ruminations on computational social science, my own learning processes and discoveries, and related technical topics and snippets.

## Education

*MS, Data Analytics and Computational Social Science* | 2022

University of Massachusetts, Amherst | Amherst, MA

*BA, Social Science* | 1992

Hampshire College | Amherst, MA

## Experience

**Gandara Center**, West Springfield, MA | Senior Systems and Analytics Engineer / CIO / IT Director | 2005 - Present

**Silicon Goblin Technologies**, Amherst, MA | Owner / Programmer | 1999 - 2015

**Rojo Networks**, San Francisco, CA | Systems Administrator and Customer Service Lead | 2006-2008

**National Center on Adult Literacy**, University of Pennsylvania Graduate School of Education, Philadelphia, PA | Systems Programmer and Lead Web Developer | 1997-1999

**Holyoke Public Schools**, Holyoke, MA | Literacy and GED Instructor | 1994-1997

**Unicom, Inc.**, Providence, RI | Systems Programmer and Technical Writer | 1983-1997

## Interests

- Data analytics, statistics, machine learning
- Software engineering
- Politics, culture and the state of the world
- Drawing, painting, guitar, motorcycles, baking, gaming